% ARCHIVED (Mega Huge Deconstruction / patch SAT narrative). See paper/archive/README.md.
% Moire Hurwitz Diophantine search — Lean-aligned narrative (HQIV_LEAN repository)
% Build: cd paper/archive/MHD && pdflatex mhd.tex
\documentclass[11pt,a4paper]{article}

\usepackage{amsmath,amssymb,amsthm,mathtools,booktabs}
\usepackage{microtype}
\usepackage[hidelinks]{hyperref}
\usepackage{xurl}
\newcommand{\lean}[1]{\path{#1}}
\usepackage[margin=1in]{geometry}
\usepackage{array}
\setlength{\emergencystretch}{3em}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\title{Moire Hurwitz Diophantine Search\\\large Mega Huge Deconstruction:\\
Moire--cusp search, arithmetic encoding, and Diophantine bookkeeping for structured witnesses}
\author{Steven Ettinger Jr}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
This note maps the \textbf{formally proved} fragments of the HQIV Lean library that touch a
\textbf{geometric SAT demo} pipeline (\texttt{scripts/hqiv\_geometric\_3sat\_demo.py}). The \textbf{only}
supported \textbf{patch search geometry} in the Lean corpus is \textbf{one-dimensional} in the patch
index~$j$: \textbf{moir\'e cumulative variation}, \textbf{cusp / threshold crossing}, and \textbf{binary
search} (\lean{Hqiv/Algebra/MoireCuspBracket.lean}, \lean{Hqiv/Algebra/MoireToyThresholdSearch.lean}).
Fourier correlations, $\Omega$-axis angles, and integer-shell embeddings feed the \textbf{same}
ordered-index score. The acronym \textbf{MHD} is \textbf{Mega Huge Deconstruction}: an explicit inventory of proved logic, analytic hooks, and open encoding bridges---without conflating BST on a monotone trace with worst-case SAT over~$2^n$
assignments.
\end{abstract}

\paragraph{Repository.}
\url{https://github.com/disregardfiat/hqiv-lean}

\paragraph{Related HQIV physics (magneto-fluid / O-Maxwell).}
The acronym \textbf{MHD} in \emph{this} document means \textbf{Mega Huge Deconstruction} (patch SAT
geometry), not magnetohydrodynamics. For the Lean \textbf{F2 chart hypothesis} tying continuum
$\varphi_F$, $\dot F$, and \texttt{delta\_theta\_prime}$(E')$ to the modified fluid vacuum source
\lean{hqivVacuumMomentumSource3}, see \path{../../HQIV_OMaxwell_fluid_chart.tex} and
\path{../../../scripts/fluid_f2_chart_alignment.py}.

\paragraph{Reader guide.}
\textbf{Green (proved):} Lean identifiers cited inline.
\textbf{Amber (scaffolds):} Fourier concentration bundles, encoding completeness for abstract~$P$.
\textbf{Red (gap):} no theorem identifies moir\'e diagnostics with \texttt{CNFFormula.eval} in general.

%--------------------------------------------------------------------
\section{Introduction: patch search = moir\'e--cusp}

Classical NP witnesses live in a huge discrete space. The HQIV pipeline separates:
\begin{enumerate}
  \item \textbf{Semantic SAT} (truth of a CNF).
  \item \textbf{Ordered patch search} along $j\in\{0,\ldots,n_{\mathrm{patch}}-1\}$ using \textbf{monotone}
        cumulative variation and BST --- the \textbf{moir\'e--cusp} layer.
\end{enumerate}
\textbf{No claim:} $\mathrm{P}=\mathrm{NP}$. \textbf{Claim in Lean:} the discrete search mechanics for
the patch are \textbf{order-theoretic} (threshold crossings, least indices).

\paragraph{``Hurwitz'' and ``Diophantine''.}
Division-algebra / Hurwitz language may appear in other HQIV scaffolds
(\lean{Hqiv/Physics/DivisionAlgebraZetaScaffold.lean}). \textbf{Diophantine} here means $\Omega$-style
literal weights and shell integers (\lean{Hqiv/Logic/CNF.lean}), not Matiyasevich.

%--------------------------------------------------------------------
\section{Logic and assignment indexing (combinatorial layer)}
\label{sec:logic}

\lean{Hqiv/Logic/CNF.lean}: literals, clauses, evaluation, satisfiability, $\omega$-encoding weights.

\lean{Hqiv/Logic/CNFPatchBridge.lean}: \texttt{assignmentFromPatchIndex}; \\
\texttt{CNFFormula.satisfiable\_iff\_exists\_patch\_index} reindexes assignments by
$j\in\mathrm{Fin}(2^n)$ (Python order). \textbf{This is reindexing, not a poly-time finder.}

\lean{Hqiv/Logic/GeometricEncodingCompleteness.lean}: abstract null-pattern $P(\phi,k)$;
\texttt{encodingCompleteness\_iff\_null\_globally\_implies\_unsat}; corollaries for ``all $k$ null
$\Rightarrow$ UNSAT'' under an encoding-completeness hypothesis.

%--------------------------------------------------------------------
\section{Moir\'e--cusp: the formal patch search layer}
\label{sec:cusp}

\lean{Hqiv/Algebra/MoireCuspBracket.lean}: \texttt{moireCumulativeAbsVariation} (monotone in $j$),
split lemmas, triangle / Gaussian-lens step on second differences.

\lean{Hqiv/Algebra/MoireToyThresholdSearch.lean}: upward closure of ``$\mathrm{cum}[j]\ge T$'',
existence of a least index above threshold --- the \textbf{BST} interface. \textbf{Explicit:} no link
to SAT in this file.

\lean{Hqiv/Algebra/MoireJerkSphereModeBridge.lean}: discrete jerk of a patch score; closed forms for
sinusoid models; ties $\Omega$ to jerk factors when the score is the intrinsic-axis sinusoid. Module doc
now states that \textbf{patch search} is the moir\'e--cusp story, not spatial vector search.

\paragraph{Conclusion (proved).}
Along the patch, the **search** object is **one-dimensional** and **order-theoretic** in~$j$; outer
$k$-sweeps are indexed diagnostics (null vs non-null), not a replacement for semantic SAT.

%--------------------------------------------------------------------
\section{Analytic hooks: $\Omega$, $\pi/(2k)$, Fourier sums (supporting the 1D score)}
\label{sec:hooks}

\lean{Hqiv/Algebra/OctonionAxisAngles.lean}: \texttt{axisAngle}, \texttt{intrinsicShellAxisAngle},
\texttt{two\_mul\_axisAngle\_eq\_pi\_div\_k}.

\lean{Hqiv/Algebra/OctonionSphereFourierPatch.lean}: peak/side correlations, \texttt{KernelIsHarmonic},
\texttt{FourierPatchConcentration}, discrete roots-of-unity sums. \textbf{Character orthogonality}
here means Fourier / DFT orthogonality on $\mathbb{Z}/n\mathbb{Z}$ --- not the removed ray geometry.

These supply **phase and spectral** structure for the **same** real score sampled along $j$. They do
\textbf{not} assert a second independent ``search direction'' in~$\mathbb{R}^8$ for the patch pipeline.

%--------------------------------------------------------------------
\section{Synthesis and gaps}
\label{sec:synth}

\textbf{Encoding completeness} for a concrete $P$ instantiated from moir\'e diagnostics remains a
\textbf{paper / analysis target}. \textbf{Target~0} (toy vs.\ Lean Fourier score) remains open; see
\texttt{AGENTS/FT\_PATCH\_CLOSED\_TARGET.md}.

\begin{center}
\begin{tabular}{@{}p{0.30\linewidth}p{0.65\linewidth}@{}}
\toprule
\textbf{Topic} & \textbf{Status} \\
\midrule
Moir\'e--cusp / BST on $j$
& Proved order theory (\lean{MoireCuspBracket}, \lean{MoireToyThresholdSearch}). \\
SAT $\leftrightarrow$ score
& Not proved. \\
Quantum ray / 2D slice search
& \textbf{Removed} from Lean (\lean{OctonionPatchQuantum} deleted). \\
\bottomrule
\end{tabular}
\end{center}

\paragraph{Acknowledgement.}
Rebuild: \texttt{lake build HQIVLEAN}.

\end{document}
